---
title: Exploratory Analysis 
author: ~
date: '2017-07-18'
slug: exploratory-analysis
categories: []
tags: []
weight: 0
---


```{r load myData, include=FALSE}
library(rvest)
library(dplyr)
library(stringr)
library(ggplot2)
load("public/data/forbes_environment.RData")
```



```{r DataMining, eval=FALSE, include=FALSE}
# Are large companies the top companies?
# 
# * What is top 
# 
# - Annual average employees
# - Aggregate employees vs Rank, assuming similar context (past five years)
# Train model: rank -> employees 


# Highlights miau --------------------------------------------------------------
# 
# dont have Toshiba's profit
# got NA's profit change and revchange


# to list data frames in environment
ls()[sapply(ls(), function(x) class(get(x))) == 'data.frame']

library(rvest)
library(dplyr)
library(stringr)
library(ggplot2)

# Load environment
load("public/data/forbes_environment.RData")

# FACTORS: Rev Change, Profits, Profit Change, Assets, Employees
# extracting each factor from its own site, ranking from employees'site


# Employees & Ranking 2015 --------------------------------------------------------------------

# Companies
employees2015.html=  read_html("/Users/iXerperience/Documents/Fernanda/blog/content/post/forbes500/Data/employees2015.html")
companies2015.span = employees2015.html %>% html_nodes("span.company-name")
companies2015 = companies2015.span %>% html_text()

# Employees
employees2015 = employees2015.html %>% 
                html_nodes("div.company-list-sort-type-container > div > div") %>%
                html_text()

# Ranking
rankings2015 = employees2015.html %>% 
  html_nodes("span.ranking") %>%
  html_text()%>% 
  as.numeric()

# Clean and convert
employees2015.numeric = str_replace_all(employees2015, "\n", "") %>%
                        str_replace_all("\t", "") %>%
                        str_replace_all(",", "")  %>% 
                        as.numeric()

# Build data frame
emp_ran.df = data.frame(ranking = rankings2015, company = companies2015, num_employees = employees2015.numeric)

# Split num_employees by tens of thousands

# emp_ran.df$group_by_employees = emp_ran.df$num_employees%/%10000


# Profits 2015 ------------------------------------------------------------
# Companies
profits2015 =  read_html("/Users/iXerperience/Documents/Fernanda/blog/content/post/forbes500/Data/profits2015.html")
companies2015profit.span = profits2015 %>% html_nodes("span.company-name")
companies2015profit = companies2015profit.span %>% html_text()


# Profits (Millions)
profits2015.text = profits2015 %>% 
  html_nodes(".sort-value") %>%
  html_text()

# Clean and convert

profits2015.numeric = str_replace_all(profits2015.text, "\n", "") %>%
  str_replace_all("\t", "") %>%
  str_replace_all(",", "")  %>% 
  as.numeric()


# Build data frame
profits.df = data.frame(company = companies2015profit, profits = profits2015.numeric)



# Merge dataframes --------------------------------------------------------
# Missing:  Rev Change,  Profit Change, Assets

# Rev Change 2015  --------------------------------------------------------
# Companies
revchange2015 =  read_html("/Users/iXerperience/Documents/Fernanda/blog/content/post/forbes500/Data/revchange2015.html")
companies2015revchange.span = revchange2015 %>% html_nodes("span.company-name")
companies2015revchange = companies2015revchange.span %>% html_text()


# Rev Change (%)
revchange2015.text = revchange2015 %>% 
  html_nodes("div.company-list-sort-type-container > div > div") %>%
  html_text()


# Cleqn and convert
revchange2015.numeric = str_replace_all(revchange2015.text, "\n", "") %>%
  str_replace_all("\t", "") %>%
  str_replace_all(",", "")  %>% 
  str_replace_all("%", "")  %>% 
  as.numeric()


# Build data frame
# miau: got NA's
revchange.df = data.frame(company = companies2015revchange, revchange = revchange2015.numeric)


# Profit Change -----------------------------------------------------------
# Companies
profitchange2015 =  read_html("/Users/iXerperience/Documents/Fernanda/blog/content/post/forbes500/Data/profitchange2015.html")
companies2015profitchange.span = profitchange2015 %>% html_nodes("span.company-name")
companies2015profitchange = companies2015profitchange.span %>% html_text()



# Profit Change (%)
profitchange2015.text = profitchange2015 %>% 
  html_nodes(".sort-value") %>%
  html_text()


# Clean and convert
profitchange2015.numeric = str_replace_all(profitchange2015.text, "\n", "") %>%
  str_replace_all("\t", "") %>%
  str_replace_all(",", "")  %>% 
  str_replace_all("%", "")  %>% 
  as.numeric()


# Build data frame
profitchange.df = data.frame(company = companies2015profitchange, profitchange = profitchange2015.numeric)



# Assets ------------------------------------------------------------------
# Companies
assets2015 =  read_html("/Users/iXerperience/Documents/Fernanda/blog/content/post/forbes500/Data/assets2015.html")
companies2015assets.span = assets2015 %>% html_nodes("span.company-name")
companies2015assets = companies2015assets.span %>% html_text()



# Assets
assets2015.text = assets2015 %>% 
  html_nodes("div.company-list-sort-type-container > div > div") %>%
  html_text()

# Clean and convert
assets2015.numeric = str_replace_all(assets2015.text, "\n", "") %>%
  str_replace_all("\t", "") %>%
  str_replace_all(",", "")  %>% 
  as.numeric()


# Build data frame
assets.df = data.frame(company = companies2015revchange, assets = assets2015.numeric)

# Complete data frame ------------------------------------------------------

fortune2015.df = merge(emp_ran.df, assets.df, by = "company") %>% 
  merge(profitchange.df, by = "company") %>%
  merge(profits.df, by = "company") %>% 
  merge(revchange.df, by = "company")

```

# Fortune's 500 companies of 2015

## Introduction and background

Google, Facebook, Apple, etc. These are some of the top companies that we frequently hear about. 

There are 497 other companies that Fortune describes as the world's 500 largest companies according to their well known ranking *Global 500*.

I wondered, what makes them the top companies? And how is the ranking constructed?




## Data source and details


In order to explore the previous question in further detail, I looked at the ranking for the year 2015 found at [Global 500 2015](http://fortune.com/global500/2015/)

The listing comprehends the following factors:

```{r factors, echo=FALSE}
names(fortune2015.df)
``` 


where,

- assets : a company's assets in millions (USD)
- profits: profits in millions (USD)
- profit change: a company's profit change between 2014 and 2015 (%)
- revchange:  a company's revenue change between 2014 and 2015 (%)

For this analysis, I have used the following libraries:

```{r libraries}
library(rvest)
library(dplyr)
library(stringr)
library(ggplot2)
```

## Initial exploratory analysis 
<!--- Visualization --->


```{r employees}
# Employees --------------------------------------------------------------
# Few companies have + 50000 employees
ggplot(fortune2015.df, aes(x = num_employees, y = ranking)) + 
  geom_point(colour = "green", alpha = 0.5) 

# Analyze outliers: few, large #employees
ggplot(fortune2015.df, aes(x = ranking, y = num_employees)) + 
  geom_point(colour = "green", alpha = 0.5)


```

There is not a clear relationship between the ranking of a company and its number of employees.
For most top companies, the number of employees oscillate between 179 (the minimum number of employees) and 500,000 employees. Some outliers stand out, with over 1,000,000 employees, particularly:

```{r outliers_employees, echo=FALSE}
outliers = fortune2015.df[which(fortune2015.df$num_employees >1000000),][1:3]
rownames(outliers) = NULL
outliers
```


```{r assets}
ggplot(fortune2015.df, aes(x = ranking, y = assets)) + geom_point(colour = "green", alpha = 0.5)
# All over the place!

```

Once again, there is not a clear relationship between ranking and assets of a company. Half of these companies' assets lie under 91,000 million USD i.e. 91 billion USD.

Among the outliers, with more than 500 billion USD in assets, 
```{r outliers_assets, echo=FALSE}
outliers_assets = fortune2015.df[which(fortune2015.df$num_employees > 500000),] %>% select(company, ranking, assets)
rownames(outliers) = NULL
outliers_assets
```


```{r profits}

type_profit.df = fortune2015.df %>% mutate(type_profit = ifelse(fortune2015.df$profits > 0, "positive", "negative"))

ggplot(type_profit.df, aes(x = ranking, y = profits, color = type_profit)) + geom_point(alpha = 0.5)  + 
  ylim(c(-10000,40000)) + geom_smooth(aes(group = 1))
# Lower the ranking, higher the profits


```


```{r profit_change}
# conglomeration near y axis
ggplot(fortune2015.df, aes(x = profitchange, y = ranking)) + geom_point(colour = "green", alpha = 0.5) + xlim(c(-500,500))
# color by positive, negative
```

As seen from the graph, there is a clearer relationship between the ranking and the profits a company. Companies at the top of the list's ranking tend to have higher profits.
Another interesting finding is that there are companies on the list which have negative profits i.e. losses.


```{r revenue_change}
type_revchange.df = fortune2015.df %>% mutate(type_revchange = ifelse(fortune2015.df$revchange > 0, "positive", "negative"))

ggplot(type_revchange.df, aes(x = ranking, y = revchange, color = type_revchange)) + geom_point(alpha = 0.5) 
```
Around half of the top companies had positive revenue change (%) between 2014 and 2015.

There were 3 companies with a significantly high revenue change (%):

```{r outliers_revchange}
outliers_revchange = fortune2015.df[which(fortune2015.df$revchange >100),] %>% select(company, ranking, revchange)
rownames(outliers_revchange) = NULL
outliers_revchange

```

## Hypothesis

From the visualization results, I presumed the factor that plays the largest role in ranking is profit. 

I also suspect that a combination of factors could influence ranking e.g. high revenue change and high profits altogether may be a key indicator of a higher position in ranking.


## Experiment / Analysis

In order to train a classification model, I have splitted the ranking positions from 1-500 to ranking buckets 1-5, where:

| Position | Bucket |
|----------|--------|
| 1-100    | 1      |
| 101-200  | 2      |
| 201-300  | 3      |
| 301-400  | 4      |
| 401-500  | 5      |



```{r buckets}
fortune2015_buckets.df = fortune2015.df %>% mutate(bucket = ceiling(fortune2015.df$ranking/100))
```


My intention in training a model was to find the most important factors to determine a company's bucket and be able to make predictions. After some googling, I noticed this multinomial log-linear model via neural networks could be useful.

First I splitted the data into training and testing, (75% to 25%).


```{r neural_network}
library(nnet)
smp_size <- floor(0.75 * nrow(fortune2015_buckets.df))

## set the seed to make your partition reproductible
set.seed(123)
train_ind <- sample(seq_len(nrow(fortune2015_buckets.df)), size = smp_size)

train <- fortune2015_buckets.df[train_ind, ]
test <- fortune2015_buckets.df[-train_ind, ]

# split training and testing data
fit <- multinom(bucket ~ num_employees + assets + profits + profitchange +revchange, train)
predict(fit, test)


predictMNL <- function(model, newdata) {
 
  # Only works for neural network models
  if (is.element("nnet",class(model))) {
    # Calculate the individual and cumulative probabilities
    probs <- predict(model,newdata,"probs")
    cum.probs <- t(apply(probs,1,cumsum))
 
    # Draw random values
    vals <- runif(nrow(newdata))
 
    # Join cumulative probabilities and random draws
    tmp <- cbind(cum.probs,vals)
 
    # For each row, get choice index.
    k <- ncol(probs)
    ids <- 1 + apply(tmp,1,function(x) length(which(x[1:k] < x[k+1])))
 
    # Return the values
    return(ids)
  }
}
predicted = predictMNL(fit,test)

# accuracy
length(which(predicted == test$bucket) == TRUE)/nrow(test)
```




```{r linear_model}
# linear model between the variables with the largest effect: ranking and profits




```


```{r decision_tree}
library(rpart)



```




}



## Results




