---
title: Exploratory Analysis 
author: ~
date: '2017-07-18'
slug: exploratory-analysis
categories: []
tags: []
weight: 0
---

```{r DataMining, include = FALSE}
# Are large companies the top companies?
# 
# * What is top 
# 
# - Annual average employees
# - Aggregate employees vs Rank, assuming similar context (past five years)
# Train model: rank -> employees 


# Highlights miau --------------------------------------------------------------
# 
# dont have Toshiba's profit
# got NA's profit change and revchange


# to list data frames in environment
ls()[sapply(ls(), function(x) class(get(x))) == 'data.frame']

library(rvest)
library(dplyr)
library(stringr)
library(ggplot2)


# FACTORS: Rev Change, Profits, Profit Change, Assets, Employees
# extracting each factor from its own site, ranking from employees'site


# Employees & Ranking 2015 --------------------------------------------------------------------

# Companies
employees2015.html=  read_html("/Users/fernandapalacios/Documents/ixperience/Week 3/blog/content/post/forbes500/Data/employees2015.html")
companies2015.span = employees2015.html %>% html_nodes("span.company-name")
companies2015 = companies2015.span %>% html_text()

# Employees
employees2015 = employees2015.html %>% 
                html_nodes("div.company-list-sort-type-container > div > div") %>%
                html_text()

# Ranking
rankings2015 = employees2015.html %>% 
  html_nodes("span.ranking") %>%
  html_text()%>% 
  as.numeric()

# Clean and convert
employees2015.numeric = str_replace_all(employees2015, "\n", "") %>%
                        str_replace_all("\t", "") %>%
                        str_replace_all(",", "")  %>% 
                        as.numeric()

# Build data frame
emp_ran.df = data.frame(ranking = rankings2015, company = companies2015, num_employees = employees2015.numeric)

# Split num_employees by tens of thousands

# emp_ran.df$group_by_employees = emp_ran.df$num_employees%/%10000


# Profits 2015 ------------------------------------------------------------
# Companies
profits2015 =  read_html("/Users/fernandapalacios/Documents/ixperience/Week 3/blog/content/post/forbes500/Data/profits2015.html")
companies2015profit.span = profits2015 %>% html_nodes("span.company-name")
companies2015profit = companies2015profit.span %>% html_text()


# Profits (Millions)
profits2015.text = profits2015 %>% 
  html_nodes(".sort-value") %>%
  html_text()

# Clean and convert

profits2015.numeric = str_replace_all(profits2015.text, "\n", "") %>%
  str_replace_all("\t", "") %>%
  str_replace_all(",", "")  %>% 
  as.numeric()


# Build data frame
profits.df = data.frame(company = companies2015profit, profits = profits2015.numeric)



# Merge dataframes --------------------------------------------------------
# Missing:  Rev Change,  Profit Change, Assets

# Rev Change 2015  --------------------------------------------------------
# Companies
revchange2015 =  read_html("/Users/fernandapalacios/Documents/ixperience/Week 3/blog/content/post/forbes500/Data/revchange2015.html")
companies2015revchange.span = revchange2015 %>% html_nodes("span.company-name")
companies2015revchange = companies2015revchange.span %>% html_text()


# Rev Change (%)
revchange2015.text = revchange2015 %>% 
  html_nodes("div.company-list-sort-type-container > div > div") %>%
  html_text()


# Cleqn and convert
revchange2015.numeric = str_replace_all(revchange2015.text, "\n", "") %>%
  str_replace_all("\t", "") %>%
  str_replace_all(",", "")  %>% 
  str_replace_all("%", "")  %>% 
  as.numeric()


# Build data frame
# miau: got NA's
revchange.df = data.frame(company = companies2015revchange, revchange = revchange2015.numeric)


# Profit Change -----------------------------------------------------------
# Companies
profitchange2015 =  read_html("/Users/fernandapalacios/Documents/ixperience/Week 3/blog/content/post/forbes500/Data/profitchange2015.html")
companies2015profitchange.span = profitchange2015 %>% html_nodes("span.company-name")
companies2015profitchange = companies2015profitchange.span %>% html_text()



# Profit Change (%)
profitchange2015.text = profitchange2015 %>% 
  html_nodes(".sort-value") %>%
  html_text()


# Clean and convert
profitchange2015.numeric = str_replace_all(profitchange2015.text, "\n", "") %>%
  str_replace_all("\t", "") %>%
  str_replace_all(",", "")  %>% 
  str_replace_all("%", "")  %>% 
  as.numeric()


# Build data frame
profitchange.df = data.frame(company = companies2015profitchange, profitchange = profitchange2015.numeric)



# Assets ------------------------------------------------------------------
# Companies
assets2015 =  read_html("/Users/fernandapalacios/Documents/ixperience/Week 3/blog/content/post/forbes500/Data/assets2015.html")
companies2015assets.span = assets2015 %>% html_nodes("span.company-name")
companies2015assets = companies2015assets.span %>% html_text()



# Assets
assets2015.text = assets2015 %>% 
  html_nodes("div.company-list-sort-type-container > div > div") %>%
  html_text()

# Clean and convert
assets2015.numeric = str_replace_all(assets2015.text, "\n", "") %>%
  str_replace_all("\t", "") %>%
  str_replace_all(",", "")  %>% 
  as.numeric()


# Build data frame
assets.df = data.frame(company = companies2015revchange, assets = assets2015.numeric)

# Complete data frame ------------------------------------------------------

fortune2015.df = merge(emp_ran.df, assets.df, by = "company") %>% 
  merge(profitchange.df, by = "company") %>%
  merge(profits.df, by = "company") %>% 
  merge(revchange.df, by = "company")

```

# Fortune's 500 companies of 2015

## Introduction and background

Google, Facebook, Apple, etc. These are some of the top companies that we frequently hear about. 

There are 497 other companies that Fortune describes as the world's 500 largest companies according to their well known ranking *Global 500*.

I wondered, what makes them the top companies? And how is the ranking constructed?




## Data source and details


In order to explore the previous question in further detail, I looked at the ranking for the year 2015 found at [Global 500 2015](http://fortune.com/global500/2015/)

The listing comprehends the following factors:

```{r factors, echo=FALSE}
names(fortune2015.df)
``` 


where,

- assets : a company's assets in millions (USD)
- profits: profits in millions (USD)
- profit change: profit change (%)
- revchange:  revenue change (%)

For this analysis, I have used the following libraries:

```{r libraries}
library(rvest)
library(dplyr)
library(stringr)
library(ggplot2)
```

## Initial exploratory analysis 
<-- Visualization -->


```{r employees}
# Employees --------------------------------------------------------------
# Few companies have + 50000 employees
ggplot(fortune2015.df, aes(x = num_employees, y = ranking)) + 
  geom_point(colour = "green", alpha = 0.5) 

# Analyze outliers: few, large #employees

ggplot(fortune2015.df, aes(x = num_employees, y = ranking)) + 
  geom_point(colour = "green", alpha = 0.5) + ylim(c(0,20))
```


```{r assets}
# Few companies have + 50000 employees
ggplot(fortune2015.df, aes(x = assets, y = ranking)) + geom_point(colour = "green", alpha = 0.5) + xlim(c(0,50000))
# All over the place!

```


```{r profits}

ggplot(fortune2015.df, aes(x = profits, y = ranking)) + geom_point(colour = "green", alpha = 0.5)  + xlim(c(-10000,40000))
# Lower the ranking, higher the profits


```


```{r profit_change}
# conglomeration near y axis
ggplot(fortune2015.df, aes(x = profitchange, y = ranking)) + geom_point(colour = "green", alpha = 0.5) + xlim(c(-500,500))
# color by positive, negative
```


```{r revenue_change}
ggplot(fortune2015.df, aes(x = revchange, y = ranking)) + geom_point(colour = "green", alpha = 0.5) + xlim(c(-50,50))
# All over
```

## Hypothesis





## Experiment / Analysis


## Results


